```{r}

```

# Study 2: Gender bias and grammar

The findings in Study 1 are consistent with both the language-as-causal and
language-as-reflection hypotheses. In Study 2, we try to distinguish between the two hypotheses by asking whether there is a relationship between psychological gender bias and language along a linguistic dimension that is unlikely to be a subject of rapid change---namely, grammatical gender. While of course grammars do change over time, they are
less malleable than the meanings of individual words, and thus less likely to be
affected by psychological biases. We predict, therefore, that if
language causally influences psychological gender biases, languages that
encode gender grammatically will tend to have larger psychological
gender biases.

### Method

We coded each of the  languages in our sample (Study 1) for grammatical gender. We used a coarse binary coding scheme, categorizing a language as encoding
grammatical gender if it made any gender distinction on noun classes
(masculine, feminine, common or neuter), and as not encoding gender
grammatically otherwise. 


### Results and Discussion

```{r}

LANGKEY_PATH <- here("data/study1b/country_langiso_langwiki_key.csv")
lang_key <- read_csv(LANGKEY_PATH) %>%
  distinct(wiki_language_code, language_name2) %>%
  rename(language_name = language_name2) %>%
  mutate(language_name = tolower(language_name))

GRAMMAR_PATH <- here("data/study2/continuous_gender_by_lang.csv")
grammar_data <- read_csv(GRAMMAR_PATH) %>%
  rename(language_name = language) %>%
  left_join(lang_key) %>%
  rename(language_code = "wiki_language_code") %>%
  mutate(empirical_binary_grammatical_gender = ifelse(empirical_grammatical_gender_score > 0, "empirically_gendered", "empirically_not_gendered"))

LANG_IAT_PATH <- here("data/study1b/iat_es_lang.csv")
iat_lang_es <- read_csv(LANG_IAT_PATH)

BEHAVIORAL_IAT_PATH <- here("data/study0/processed/by_language_df.csv")
iat_behavioral_es <- read_csv(BEHAVIORAL_IAT_PATH) %>%
  rename(language_code = "wiki_language_code") %>%
  select(language_code, median_country_age, 
         prop_male,log_age, es_iat_sex_age_order_explicit_resid,
         es_iat_sex_age_order_implicit_resid, per_women_stem, n_participants)

theoretical_gender <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/IATLANG/exploratory_analyses/10_grammatical_gender/gender_grammar.csv") %>%
  select(language_code, wikipedia_grammar_type)

all_es <- left_join(iat_behavioral_es, iat_lang_es, by = "language_code") %>%
  left_join(grammar_data) %>%
  left_join(theoretical_gender) 

```

```{r, eval = F}
ggplot(all_es , aes(y = es_iat_sex_age_order_implicit_resid, 
                    x = empirical_grammatical_gender_score)) +
  geom_label(aes(label= language_code)) +
  geom_smooth(method = "lm") +
  theme_classic()

ggplot(all_es , aes(y = lang_es_wiki, 
                    x = empirical_grammatical_gender_score)) +
  geom_label(aes(label= language_code)) +
  geom_smooth(method = "lm") +
  theme_classic()

t.test(es_iat_sex_age_order_implicit_resid ~ empirical_binary_grammatical_gender + , 
       data = all_es,
       paired = F)

t.test(lang_es_wiki ~ empirical_binary_grammatical_gender, 
       data = all_es,
       paired = F)

lm(es_iat_sex_age_order_implicit_resid ~ empirical_binary_grammatical_gender + median_country_age, 
       data = all_es) %>%
  summary()

lm(es_iat_sex_age_order_explicit_resid ~ empirical_binary_grammatical_gender + median_country_age, 
       data = all_es) %>%
  summary()


```




 