---
title: Study 1C analyses
subtitle: ED6
date: "`r Sys.Date()`"
output: latex_document
---


```{r setup, include = F}
# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(here)
opts_chunk$set(echo = F, message = F, warning = F,
               error = F, tidy = F, cache = F)
```

```{r}
full_width = 7.08661
FIG_OUTPATH <- here("writeup/journal/supporting_information/ED/ED7/ED7a.pdf")

```


```{r}
DATA_A <- here("writeup/journal/supporting_information/ED/ED7/ED7_data_1.csv")
DATA_B <- here("writeup/journal/supporting_information/ED/ED7/ED7_study_1c_model_print.txt")
```


## Plot


```{r, fig.height = 3.6}
study1c_data <- read_csv(DATA_A) %>%
  mutate(group = case_when(data_type == "confirmatory" & corr == "Behavioral Bias Difference ~ Language Bias Difference" ~ "targ", TRUE ~ "other"))

p1 <- ggplot(study1c_data, aes(x = corr, y = mean, color = group)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  facet_grid(~ fct_rev(data_type)) +
  geom_pointrange(aes(ymin = ci_lower, max = ci_upper)) +
  theme_classic(base_size = 7) +
  scale_color_manual(values = c("black", "red")) +
  xlab("Correlation Type") +
  ylab("Mean correlation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
          plot.margin = margin(0, 0, 0, 5, "cm"))
p1
```

## Mixed-effect model

```{r}
model_print <- readChar(DATA_B, file.info(DATA_B)$size)
```

\begin{table}[H]
\centering
\begin{tabular}{|l|l|l|l|}
\hline

& \bf{Behavioral Effect Resid} & & \\ \hline
\bf{Predictors} & Estimates & SE & Statistic \\ \hline
(intercept) & -0.00 & 0.05 & -0.09 \\ \hline
country (uk) & 0.02 & 0.01 & 1.74 \\ \hline
language bias difference (uk - us) & -0.03 & 0.07 & -0.42 \\ \hline
country:language bias difference & 0.05 & 0.02 & 2.88 \\ \hline
\bf{Random Effects} \\ \hline
s2 & 0.20  & & \\ \hline
T00 user id & 0.01  & & \\ \hline
T00 domain & 0.07 & &  \\ \hline
N user id & 22059  & & \\ \hline
N domain & 31  & & \\ \hline
\specialrule{.1em}{.05em}{.05em} 
Observations & 27045  & & \\ \hline

\end{tabular}
\end{table}


```{r}
pdf(FIG_OUTPATH, width = full_width, height = 4.2)
p1
dev.off()
```

## Title and Legend 


The exact pre-registered analysis (https://osf.io/3f9ed/) of Study 1c is presented below. Pairwise correlations between all variables (language bias, behavioral bias, and UK-US difference measures) are shown, averaging across estimates of language bias from the 5 model runs. Error bars are 95% CIs. As stated in the pre-registration, the key test of our hypothesis is that the correlation between the UK - US linguistic difference ("Language Bias Difference") and the UK - US behavioral difference  ("Behavioral Bias Difference") is greater than 0 (shown in red below). That data are consistent with this prediction.

The confirmatory dataset is shown on the right, along with the smaller exploratory dataset on the left for reference. 

The full results to the mixed-effect model described in the paper are presented below:
