---
title: Comparing behavior vs. language
author: Molly Lewis 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: no
    toc: yes
---
  
```{r setup, include = F}

# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(langcog)
library(stringr)
library(forcats)
library(broom)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)

```
Here we look at whether IAT scores correlate with language bias for a range of difference languages. A number of different IAT measures are examined. Note this is at the language level (averaging across countries that speak the same language for the IAT data).


Read in behavior and language biases
```{r}
countries_langs <- read_csv("../../data/other/countries_lang.csv") %>%
  mutate(language_name = ifelse(language_name == "Spanish; Castilian", "Spanish", language_name),
         language_name = ifelse(language_name == "Dutch; Flemish", "Dutch", language_name))

IAT_behavior_measures <- read_csv("../behavior_IAT/IAT_behavior_measures.csv") %>%
  left_join(countries_langs %>% select(country_name, language_name, language_code)) %>%
  filter(type %in% c("country_means_D_score", 
             "country_gender_D_score", 
             "country_gender_D_score_diff", 
             "country_RT_mean",
             "country_RT_ratio")) %>%
  group_by(type, language_code, sex) %>%
  summarize(mean = mean(mean)) 

lang_bias <- read_csv("career_effect_sizes.csv") 

lang_bias_google <- read_csv("google_translate/career_effect_sizes_google.csv",
                             col_names = c("language_code",   "test",  "test_name", "lang_bias")) %>%
  filter(language_code != "he", language_code != "zu", language_code != "th") #,language_code !=  "ha",language_code !=  "id",language_code !=  "zh")


d <- left_join(IAT_behavior_measures, lang_bias) %>%
  mutate_if(is.character, as.factor) %>%
  filter(!is.na(lang_bias))
```

There are `r length(unique(d$language_code))` languages represented here.


## Measures
### Behavior
```{r, fig.width = 5, fig.heigh = 5}
ggplot(d, aes(x = mean)) +
  geom_histogram() +
  facet_wrap(~type, scales = "free_x")+
  theme_bw()
```

### Language - Hand translations
```{r, fig.height = 3, fig.width = 3}
ggplot(lang_bias, aes(x = lang_bias)) +
  geom_histogram() +
  theme_bw()
```

### Language - google translations
```{r, fig.height = 3, fig.width = 3}
ggplot(lang_bias_google, aes(x = lang_bias)) +
  geom_histogram() +
  theme_bw()
```

How correlated are the two language measures?
```{r}
left_join(lang_bias_google, lang_bias, by = "language_code") %>%
  do(tidy(cor.test(.$lang_bias.x, .$lang_bias.y))) %>%
  kable()
```

## Correlation between measures - Hand-translations

```{r,  fig.width = 10, fig.height = 2.5}
d %>%
  mutate(type2 = ifelse(type %in% c("country_gender_D_score_diff", "country_means_D_score", "country_gender_D_score"), "D-score", ifelse(type %in% c("country_RT_ratio", "country_RT_mean"), "RT", "explicit"))) %>%
ggplot( aes(x = mean, y = lang_bias, group = sex, color = sex, shape = type2)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(.~type2 + type, scales = "free_x", drop = T) +
  theme_bw()
```

Two effects:
(1) D-score correlated with language bias (driven by males)
(2) Explicit family-male associaiton correlated with language bias

```{r, fig.width = 10}
d %>%
  group_by(type, sex) %>%
  do(tidy(cor.test(.$mean, .$lang_bias))) %>%
  arrange(p.value) %>%
  select(-parameter, -method, -alternative) %>%
  kable()
```

## Correlation between measures - Google-translations
```{r}
d <- left_join(IAT_behavior_measures, lang_bias_google) %>%
  mutate_if(is.character, as.factor) %>%
  filter(!is.na(lang_bias))

```

There are `r length(unique(d$language_code))` languages represented here.

```{r,  fig.width = 10, fig.height = 2.5}
d %>%
  mutate(type2 = ifelse(type %in% c("country_gender_D_score_diff", "country_means_D_score", "country_gender_D_score"), "D-score", ifelse(type %in% c("country_RT_ratio", "country_RT_mean"), "RT", "explicit"))) %>%
ggplot( aes(x = mean, y = lang_bias, 
            group = sex, color = sex, shape = type2)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(.~type2 + type, scales = "free_x", drop = T) +
  theme_bw()
```

```{r, fig.width = 10}
d %>%
  group_by(type, sex) %>%
  do(tidy(cor.test(.$mean, .$lang_bias))) %>%
  arrange(p.value) %>%
  select(-parameter, -method, -alternative) %>%
  kable()
```

