---
title             : "Language use shapes cultural stereotypes: Large scale evidence from gender"
shorttitle        : "Language use shapes cultural stereotypes"
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{t} #make every figure with caption = t

author: 
  - name          : "Molly Lewis"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "mollyllewis@gmail.com"
    address       : ""
  - name          : "Gary Lupyan"
    affiliation   : "1"
    
affiliation: 
  - id            : "1"
    institution   : "University of Wisconsin-Madison"
  - id            : "2"
    institution   : "University of Chicago"

authornote: |
    Portions of this manuscript appeared in Lewis & Lupyan, 2018, Cog. Sci Proceedings. 

abstract: | 
  Cultural stereotypes such as the idea that men are more suited for paid work while women for taking care of the home and family may contribute to gender imbalances in STEM fields (e.g., Leslie, Cimpian, Meyer, & Freeland, 2015) and other undesirable gender disparities. Here, we test the hypothesis that word co-occurrence statistics (e.g., the co-occurrence of “nurse” with “she”) play a causal role in the formation of the men-career/women-family stereotype. We use word embedding models to measure bias in the distributional statistics of 25 languages and find that languages with larger biases tend to have speakers with larger implicit biases (_N_ = 657,335). These biases are further related to the extent that languages mark gender in their lexical forms (e.g., “waiter”/“waitress”) hinting that linguistic biases may be causally related to biases shown in people's implicit judgments.
  
keywords          : "cultural stereotypes, implicit association task (IAT), gender"
wordcount         : ""

bibliography      : ["library.bib"]

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library(papaja)
library(kableExtra)
library(here)
source(here("writeup/journal/helpers/psych_to_mat.R"))
```

```{r analysis-preferences}
set.seed(42) # Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, 
                      warning = F, 
                      cache = F, 
                      fig.pos = "t!")
```

# Introduction
By the time they are two years old, children have begun to acquire the gender stereotypes in their culture [@gelman2004mother]. These stereotypes can have undesirable real-world consequences. For example, in one study, girls, compared to boys, were less likely to think that girls are "brilliant" and also less likely to choose activities that were described as for children "who are very, very smart"  [@bian2017gender]. In the aggregate, these behavioral choices could lead to the observed lower rates of female participation in science, technology, engineering, and mathematics (STEM) fields [@miller2015women; @stoet2018gender; @ceci2011understanding; @leslie2015expectations]. Given the potential downstream consequences of cultural stereotypes, it is important to understand how cultural stereotypes are formed.

We can distinguish between two major sources of information on which gender stereotypes may be based. One is direct experience. For example, one may observe that most nurses are women and most philosophers are men and conclude that women are better suited for nursing and men for philosophy. Another source of information is language. Even in the absence of any direct experience with nurses or philosophers, a learner may observe that language about nurses and philosophers is more associated with female and male contexts, respectively. These contexts include  gender-specific proper names (e.g., "Sam" vs. "Ashley"), pronouns (e.g., "he" vs. "she"), titles (e.g., "waiter" vs. "waitress"), and higher-order associations (e.g., seemingly non-gendered words can become gendered by being associated with explicitly gendered contexts). An additional linguistic  source of information about gender is morphological markers in languages with sex-based grammatical gender systems (approximately 32% of languages; Dryer \& Haspelmath, 2013). For example, in Spanish the gender of a nurse must be specified grammatically ("enfermer*a*" vs. "enfermer*o*"). In all of these cases, the presence of linguistic cues to gender may exaggerate the extent to which gender information is communicated to listeners.

Past research shows that even young children are sensitive to gender information delivered through language. For example, young children perform worse in a game if they are told that an anonymous member of the opposite gender performed better than they did on a previous round [@rhodes2008preschoolers], or merely told that the game is associated with a particular gender [@cimpian2012good]. Further, there is evidence that descriptions as minimal  as a single word can influence children's stereotypes: In one study, children were more likely to infer that a novel skill is stereotypical of a gender  if the skill was introduced to children with a generic as opposed  a non-generic subject ["[Girls are/There is a girl who is] really good at a game called 'gorp' "; @cimpian2011generic]. To the extent that language is a source of information for forming cultural stereotypes, two people with similar direct experiences, but different linguistic experiences, may come to have different stereotypes.

A common way of quantifying cultural stereotypes in individuals is by using the _Implicit Association Test_ [IAT; @greenwald1998measuring] which measures differences in reaction time in associations between different concepts (e.g. male-related words and career-related words vs. family-related words). As it turns out, various biases studied using IATs can be predicted from the distributional statistics of language (word co-occurrences). Using these statistics, Caliskan, Bryson, and Narayanan (2017; henceforth _CBN_) measured the  distance between the words presented to participants in the IAT task. CBN found that these distances were highly correlated with the biases computed by a variety of IATs (e.g., valence and Caucasian vs. African-American names; gender and math vs. arts; permanence and mental vs. physical diseases). CBN only measured semantic biases in English. Here, we extend CBN’s method to 25 languages examining whether languages with a stronger gender bias as expressed in distributional semantics predict stronger implicit and explicit gender biases on a large dataset of previously administered gender-career IAT (_N_ = 657,335; Nosek, et al., 2002). 

Discovering that stronger biases in language correlate with stronger biases in people’s behavior can be interpreted in at least two ways. The first is that language merely _reflects_ people’s biases which are learned chiefly through nonlinguistic experiences. We refer to this as the _language as reflection_ hypothesis. The second possibility is that language exerts a causal influence on people’s biases. We refer to this as the _language as causal factor_ hypothesis. Work showing that linguistic descriptions can impact stereotypes in-the-moment [@rhodes2008preschoolers; @cimpian2012good; @cimpian2011generic] already suggests that language _can_ have some causal impact in an experimental context. We were interested in whether it actually does, and by what means.


In Study 1a, we examine whether different languages convey gender roles differently in their distributional structure. That is, our analysis is not concerned with _what_ is being said about typical gender roles, but rather the ways that _potentially_ gendered words may be gendered by the morphological rules and statistical structure of different languages. In Study 1b we use this information to predict responses on a gender-career IAT. In Study 2 we examine whether the extent to which different languages have grammatical gender systems and use different forms for occupation terms (e.g., "waiter"/"waitress" but "teacher"/"teacher") correlates with greater implicit gender bias thereby helping to narrow down the source of linguistic knowledge that may be playing a role in shaping gender stereotypes. Together, our data suggest that language not only reflects existing biases, but likely plays a causal role in shaping culturally-specific notions of gender.


```{r child = "sections/study0_writeup.Rmd"}
```

```{r child = "sections/study1_writeup.Rmd"}
```

```{r child = "sections/study2_writeup.Rmd"}
```

# General Discussion
Across two studies, we sought to understand the role that language plays in the formation of one particular gender stereotype: the bias to associate men with career and women with family. We estimated the magnitude of this bias among speakers of 25 different languages using a large scale administration of the IAT, and then examined how the strength of the stereotype related to gender biases found in language. In Study 1, we found that languages that have a higher degree of gender bias in their co-occurrence statistics tend to have speakers that have stronger gender stereotypes. In Study 2, we found that words that were explicitly marked for the gender of the referent ("waiter"/"waitress") tended to have more biased statistics, and that languages that more frequently  explicitly marked the gender of referents tended to have speakers that were overall more biased. 

Our work is the first to characterize the relationship between broad structural patterns in language and cultural stereotypes. Broadly, the positive correlation that we find between these two variables -- gender bias in language and gender bias in speakers -- is consistent with both the possibility that language plays a causal role in the emergence of cultural stereotypes and the idea that language merely reflects existing stereotypes in its speakers. Study 2 provides suggestive evidence for the causal hypothesis: The positive correlation between the degree of gender-specific gender labels in a language and implicit psychological bias is most parsimoniously explained by the language-as-causal-hypothesis, since it is unlikely that language forms change on a timescale that could directly reflect behavior. Importantly, however, these two causal forces are not mutually exclusive, and in fact may amplify the effects of each other.  Future work could use experimental methods to  manipulate language statistics in order to more directly examine these causal influences.

An important limitation of the current work is that our measure of the psychological reality of the gender stereotype relies on the IAT, a measure that has been criticized for both its low reliability [@lane2007understanding] and limited validity [e.g., @oswald2013predicting]. Issues of reliability are less relevant here because we use the IAT to measure group-level differences, rather than as an individual-difference measure. However, issues of validity are an important concern, particularly since we find that language estimates of bias are uncorrelated with explicit measures, although these measures were extremely coarse [others have pointed to IAT's lack of predictive validity; @fazio2003implicit]. Understanding the full import of linguistic biases on cultural stereotypes would therefore require obtaining measures more closely related to real-world behavior.

In conclusion, information about cultural stereotypes is necessarily acquired through experience in that culture. In our work here, we have suggested that the statistics of language use are an important element of that experience: The mere process of listening to and producing language exposes one to statistics that may lead to the formation of cultural stereotypes.  Many cultural associations and stereotypes are present in the statistics of language may be innocuous -- indeed, these statistics may be an important mechanism through which cultural information is transmitted. But, in other cases, like the kind of gender stereotypes investigated here, language may be playing a powerful role in their formation, and ultimately contributing to undesirable real-world consequences like gender inequality in STEM. Understanding  the causal role that language plays in the formation of these stereotypes is therefore an important first step to changing these consequences.
 
\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

---
nocite: | 
  @caliskan2017semantics
  @ciafactbook
  @nosek
  @stoet2018gender
  @miller2015women
  @schoemann2017determining
...


\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
