---
title             : "Language use shapes cultural stereotypes: Large scale evidence from gender"
shorttitle        : "Language use shapes cultural stereotypes"

author: 
  - name          : "Molly Lewis"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "mollyllewis@gmail.com"
    address       : ""
  - name          : "Gary Lupyan"
    affiliation   : "1"
    
affiliation: 
  - id            : "1"
    institution   : "University of Wisconsin-Madison"
  - id            : "2"
    institution   : "University of Chicago"

authornote: |
    Portions of this manuscript appeared in Lewis & Lupyan, 2018, Cog. Sci Proceedings. 

abstract: | 
  this is an abstract

  
keywords          : "cultural stereotypes, implicit association task (IAT), gender"
wordcount         : "X"

bibliography      : ["library.bib"]

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library(papaja)
library(kableExtra)
library(here)
source(here("writeup/journal/helpers/psych_to_mat.R"))
```

```{r analysis-preferences}
set.seed(42) # Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, 
                      warning = F, cache = F)
```

# Introduction
By the time they are two years old, children have begun to acquire the gender stereotypes in their culture [@gelman2004mother]. These stereotypes can have undesirable real-world consequences. For example, in one study, girls, compared to boys, were less likely to think that girls are "brilliant" and also less likely to choose activities that were described as for children "who are very, very smart"  [@bian2017gender]. In the aggregate, these behavioral choices could lead to the observed lower rates of female participation in science, technology, engineering, and mathematics (STEM) fields [@miller2015women; @stoet2018gender; @ceci2011understanding; @leslie2015expectations]. Given the potential downstream consequences of cultural stereotypes, it is important to understand how the stereotypes are formed.
<!-- "where they come from" is ambiguous.. Many explanations would be historical/anthropological and that's not what we're studying here -->

We can distinguish between two major sources of information on which stereotypes may be based is linguistic information. One is direct experience. For example, one may observe that most nurses one encounters are women and most philosophers are men and conclude that women better suited for nursing and men for philosophy. Another, non-mutually exclusive source of information is linguistic. Even in the absence of any direct experience with nurses or philosophers, a learner may observe that language about nurses and philosophers is more associated with female and male contexts, respectively. These contexts include proper names, pronouns, grammatical markers (particularly for languages with grammatical gender), and gendered contexts more generally. Past research shows that even young children are sensitive to such linguistically delivered information. For example, young children perform worse in a game if they are told that an anonymous member of the opposite gender performed better than they did on a previous round [@rhodes2008preschoolers], or merely told that the game is associated with a particular gender [@cimpian2012good]. Further, there is evidence that descriptions as minimal  as a single word can influence children's stereotypes: In one study, children were more likely to infer that a novel skill is stereotypical of a gender  if the skill was introduced to children with a generic as opposed  a non-generic subject ["[Girls are/There is a girl who is] really good at a game called 'gorp' "; @cimpian2011generic]. To the extent that language is an important source of information for forming cultural stereotypes, two people with similar direct experiences, but different linguistic experiences may come to have different stereotypes.

<!-- sentence introducing IAT as a measure of stereotypes --> As it turns out, various biases studied using IATs can be predicted from language using distributional semantics which model word meanings as the contexts in which the word occurs. Caliskan, Bryson, and Narayanan (2017; henceforth _CBN_) measured the distance in vector space between the words presented to participants in the IAT task. CBN found that these distance measures were highly correlated with reaction times in the behavioral IAT task. For example,  CBN find a bias to associate males with career and females with family in the career-gender IAT, suggesting that the biases measured by the IAT are also found in the lexical semantics of natural language. CBN only measured semantic biases in English. Here, we extend CBN’s method to 26 languages examining whether languages with a stronger gender bias as expressed in distributional semantics predict stronger implicit and explicit gender biases on a large dataset of a gender-career IAT previously administered [n=XXXX; REF]. 

Discovering that stronger biases in language correlate with stronger biases in people’s behavior can be interpreted in two ways. The first is that language merely _reflects_ people’s biases which are learned chiefly through nonlinguistic experiences. We refer to this as the _language as reflection_ hypothesis. The second possibility is that language exerts a causal influence on people’s biases. We refer to this as the _language as causal agent_ hypothesis. Work showing that linguistic descriptions can impact stereotypes in-the-moment [REFS cimpian etc.] already suggests that language _can_ have some causal impact in an experimental context. We were interested in whether it actually does, and by what means.

Languages convey gender information in multiple ways including gender-specific proper names, pronouns, and titles (e.g., waiter vs. waitress). In addition, approximately 25% of languages  have some type of grammatical gender [@corbett1991] in which morphological markers are used to signal gender. <!-- The more informative numbers/references may be here.. https://wals.info/chapter/31  It’s more than 25%. Please correct as needed !-->. Some previous work has shown that gender information conveyed by such morphological markers---which are also extended to inanimate objects---can infuse the inanimate objects with “natural” gender [@phillips2003can; @sera1994grammatical]. But more importantly for present purposes, because grammatical gender markers enter into agreement patterns, their use can exaggerate the extent to which gender is being communicated. For example, in Spanish the gender of a nurse has to be signaled grammatically: enfermer_a_ vs. enfermer_o_. In Russian verbs (in some tenses)  are inflected based on the gender of the speaker: ya ustal (I am tired [m], but Ya ustala (I am tired [f]). In both cases, it is not possible to leave the gender unspecified. <!-- may want to use standard linguistic typesetting for these langauge examples rather than italics/parentheses --> 

In study 1a we examine the degree to which gender in different languages is encoded in their distributional structure. That is, our analysis is not concerned with _what_ is being said about typical gender roles, but rather the statistical patterns to which _potentially_ gendered words are gendered by different languages. In study 1b we use this information to predict previously collected responses on an gender-career IAT. In study 2 we examine whether the extent to which different languages use different forms for occupations (e.g., waiter/waitress but teacher/teacher) correlates with greater implicit gender bias thereby helping to narrow down the source of linguistic knowledge that may be playing a role in shaping gender stereotypes. Together, our data suggest that language not only reflects existing biases, but likely plays a causal role in shaping culturally-specific notions of gender.


```{r child = "sections/study0_writeup.Rmd"}
```

```{r child = "sections/study1_writeup.Rmd"}
```

```{r child = "sections/study2_writeup.Rmd"}
```

# General Discussion

* IAT is only behavioral in a weak sense
* socialization in development
* implicit vs. explicit difference
* pragmatics related to presense of distinction for kids - must be relevenat! - chesnut stuff
* gender asymmetries in the IAT (women show bigger effect than men?)
* experimental work to get at causality more so

\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

---
nocite: | 
  @caliskan2017semantics
  @ciafactbook
  @nosek
...


\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
