---
title             : "Language use shapes cultural stereotypes: Large scale evidence from gender"
shorttitle        : "Language use shapes cultural stereotypes"

author: 
  - name          : "Molly Lewis"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "mollyllewis@gmail.com"
    address       : ""
  - name          : "Gary Lupyan"
    affiliation   : "1"
    
affiliation: 
  - id            : "1"
    institution   : "University of Wisconsin-Madison"
  - id            : "2"
    institution   : "University of Chicago"

authornote: |
    Portions of this manuscript appeared in Lewis & Lupyan, 2018, Cog. Sci Proceedings. 

abstract: | 
  this is an abstract

  
keywords          : "cultural stereotypes, implicit association task (IAT), gender"
wordcount         : "X"

bibliography      : ["library.bib"]

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library(papaja)
library(kableExtra)
library(here)
source(here("writeup/journal/helpers/psych_to_mat.R"))
```

```{r analysis-preferences}
set.seed(42) # Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, 
                      warning = F, cache = F)
```

# Introduction
By the time they are two years old, children have already begun to acquire the gender stereotypes in their culture [@gelman2004mother]. This finding is  concerning given evidence that children's conceptualizations of gender can have undesirable real-world consequences. For example, in one study, girls, compared to boys, were less likely to think that girls are "brilliant" and also less likely to choose activities that were described as for "children who are very, very smart"  [@bian2017gender]. In the aggregate, these behavioral choices could lead to the observed lower rates of female participation in science, technology, engineering, and mathematics (STEM) fields [@miller2015women; @stoet2018gender; @ceci2011understanding; @leslie2015expectations]. Given the potential downstream consequences of cultural stereotypes, it is therefore important to understand their origins.

While there are likely many  factors that shape cultural stereotypes, a large body of evidence suggests that language may play an important causal role.  For instances, young children perform worse in a game if they are told that an anonymous member of the opposite gender performed better than they did on a previous round [@rhodes2008preschoolers], or merely told that the game is _associated_ with a particular gender [@cimpian2012good]. Further, there is evidence that descriptions as minimal  as a single word can influence children's stereotypes: In one study, children were more likely to infer that a novel skill is stereotypical of a gender  if the skill was introduced to children with a generic as opposed  a non-generic subject ["[Girls are/There is a girl who is] really good at a game called 'gorp' "; @cimpian2011generic].

In our work here, we  ask whether language influences psychological gender biases in ways that are more subtle than linguistic descriptions or word choice. In particular, we test two possible mechanisms through which language might shape gender biases. The first is through word meaning derived from a word's co-occurrence with other words, an approach to word meaning known as _distributional semantics_. Under this approach, words that tend to occur in similar contexts in language may lead speakers to assume---either implicitly or explicitly---that they have similar meanings. For example, statistically, the word "nurse" occurs in many of the same contexts as the pronoun "her," providing an implicit link between these two concepts that may lead to a bias to assume that nurses are female. This route may be particularly influential because the bias is encoded in language in a way that is implicit  and thus may be more difficult to discount. To our knowledge, no work to date has tested the link between distributional semantics and social stereotypes. 

The second way word meaning might shape gender biases is through the overt grammatical marking of gender, particularly on nouns, which is obligatory in roughly one quarter of languages [e.g., in Spanish, “ni&ntilde;a” (girl) and “enfermera” (nurse) both take the gender marker _-a_ to indicate grammatical femininity; @corbett1991]. Because grammatical gender has a natural link to the real world, speakers may assume that grammatical markers are meaningful even when applied to inanimate objects that do not have a biological sex.  In addition, the mere presence of obligatory marking of grammatical gender may  promote bias  by making the  dimension of gender more salient to speakers.

Past experimental work suggests a causal link between grammatical gender and psychological gender bias in both adults [e.g.,@phillips2003can] and children [e.g., @sera1994grammatical]. For example, Phillips and Boroditsky (2003) asked Spanish-English and German-English adult bilinguals to make similarity judgments between pairs of pictures depicting an object with a natural gender (e.g., a bride) and one without (e.g., a toaster). They found that participants rated pairs as more similar when the pictures matched in grammatical gender in their native language. While these types of studies provide suggestive evidence for a causal link between language and psychological gender bias, they are limited by the fact that they typically only compare speakers of  2-3 different languages and measure bias in a way that is subject to demand characteristics.

Of course, evidence for a close correspondence between language and psychological gender biases is also consistent with the simpler explanation that language reflects a pre-existing gender bias in its speakers  (_language-as-reflection hypothesis_). We assume that the language-as-reflection hypothesis is true to some extent: some of the ways we talk about gender reflect our knowledge and biases acquired independently of language. For example, we may observe that most nurses are women, and therefore be more likely to use a female pronoun to refer to a nurse of an unknown gender.  Our present goal is to understand the extent to which language may also exert a causal influence on conceptualizations of gender (_language-as-causal hypothesis_).

In what follows, we ask whether the way gender is linguistically encoded
across 26 different languages predicts cross-cultural variability in a specific manifestation of a gender bias---the bias to associate men with careers and women with family. We begin by describing our  cross-cultural dataset of psychological gender bias. We then examine whether variability in language, as captured by distributional semantics, predicts differences in the gender-career implicit bias (Study 1). We next ask whether the presence of grammatical gender in a language is associated with greater implicit gender bias (Study 2). Together, our data suggest that language not only reflects existing biases, but plays a causal role in shaping culturally-specific notions of gender.


```{r child = "sections/study0_writeup.Rmd"}
```

```{r child = "sections/study1_writeup.Rmd"}
```

```{r child = "sections/study2_writeup.Rmd"}
```

# General Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

---
nocite: | 
  @caliskan2017semantics
  @ciafactbook
  @nosek
...


\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
