---
title             : "Language use shapes cultural stereotypes: Large scale evidence from gender"
shorttitle        : "Language use shapes cultural stereotypes"

author: 
  - name          : "Molly Lewis"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "mollyllewis@gmail.com"
    address       : ""
  - name          : "Gary Lupyan"
    affiliation   : "1"
    
affiliation: 
  - id            : "1"
    institution   : "University of Wisconsin-Madison"
  - id            : "2"
    institution   : "University of Chicago"

authornote: |
    Portions of this manuscript appeared in Lewis & Lupyan, 2018, Cog. Sci Proceedings. 

abstract: | 
  People of all ages share cultural stereotypes, like the gender stereotype that men are more closely related to careers and women are more closely related to family. Importantly, this gender stereotype may contribute to the imbalance in participation between men and women in STEM fields (e.g., Leslie, Cimpian, Meyer, & Freeland, 2015), in addition to many other undesirable gender disparities. Here, we test the hypothesis that language, particularly co-occurrence statistics of language use (e.g., the frequency of "nurse" with the pronoun "she"), plays a causal role in the formation of this gender stereotype. We use word embedding models to measure the magnitude of the bias to associate men with career and women with family in the statistics of 25 different languages, and find that the magnitude of this bias is positively correlated with magnitude of psychological bias among speakers of that languages (N = 764,520), as measured by the Implicit Association Test (IAT). Further, we find that biases in language statistics are exaggerated for words that lexically mark the gender of the referent (e.g. waiter/waitress). Our work provides the first large-scale demonstration of the relationship between language and psychological biases. 

  
keywords          : "cultural stereotypes, implicit association task (IAT), gender"
wordcount         : "X"

bibliography      : ["library.bib"]

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library(papaja)
library(kableExtra)
library(here)
source(here("writeup/journal/helpers/psych_to_mat.R"))
```

```{r analysis-preferences}
set.seed(42) # Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, 
                      warning = F, cache = F)
```

# Introduction
By the time they are two years old, children have begun to acquire the gender stereotypes in their culture [@gelman2004mother]. These stereotypes can have undesirable real-world consequences. For example, in one study, girls, compared to boys, were less likely to think that girls are "brilliant" and also less likely to choose activities that were described as for children "who are very, very smart"  [@bian2017gender]. In the aggregate, these behavioral choices could lead to the observed lower rates of female participation in science, technology, engineering, and mathematics (STEM) fields [@miller2015women; @stoet2018gender; @ceci2011understanding; @leslie2015expectations]. Given the potential downstream consequences of cultural stereotypes, it is important to understand how the stereotypes are formed.

We can distinguish between two major sources of information on which stereotypes may be based is linguistic information. One is direct experience. For example, one may observe that most nurses one encounters are women and most philosophers are men and conclude that women better suited for nursing and men for philosophy. Another, non-mutually exclusive source of information is linguistic. Even in the absence of any direct experience with nurses or philosophers, a learner may observe that language about nurses and philosophers is more associated with female and male contexts, respectively. These contexts include proper names, pronouns, grammatical markers (particularly for languages with grammatical gender), and gendered contexts more generally.

Past research shows that even young children are sensitive to such linguistically delivered information. For example, young children perform worse in a game if they are told that an anonymous member of the opposite gender performed better than they did on a previous round [@rhodes2008preschoolers], or merely told that the game is associated with a particular gender [@cimpian2012good]. Further, there is evidence that descriptions as minimal  as a single word can influence children's stereotypes: In one study, children were more likely to infer that a novel skill is stereotypical of a gender  if the skill was introduced to children with a generic as opposed  a non-generic subject ["[Girls are/There is a girl who is] really good at a game called 'gorp' "; @cimpian2011generic]. To the extent that language is an important source of information for forming cultural stereotypes, two people with similar direct experiences, but different linguistic experiences may come to have different stereotypes.

To measure individuals' cultural stereotypes, researchers have commonly used a task that uses reaction time as an index of the degree of psychological association between two concepts (e.g. males and career), known as the _Implicit Association Task_ [IAT; @greenwald1998measuring]. As it turns out, various biases studied using IATs can be predicted from language using distributional semantics which model word meanings as the contexts in which the word occurs. Caliskan, Bryson, and Narayanan (2017; henceforth _CBN_) measured the distance in vector space between the words presented to participants in the IAT task. CBN found that these distance measures were highly correlated with reaction times in the behavioral IAT task. For example,  CBN find a bias to associate males with career and females with family in the career-gender IAT, suggesting that the biases measured by the IAT are also found in the lexical semantics of natural language. CBN only measured semantic biases in English. Here, we extend CBN’s method to 26 languages examining whether languages with a stronger gender bias as expressed in distributional semantics predict stronger implicit and explicit gender biases on a large dataset of a gender-career IAT previously administered (_N_ = 778,795; Nosek, Banaji, & Greenwald, 2002). 

Discovering that stronger biases in language correlate with stronger biases in people’s behavior can be interpreted in two ways. The first is that language merely _reflects_ people’s biases which are learned chiefly through nonlinguistic experiences. We refer to this as the _language as reflection_ hypothesis. The second possibility is that language exerts a causal influence on people’s biases. We refer to this as the _language as causal agent_ hypothesis. Work showing that linguistic descriptions can impact stereotypes in-the-moment [@rhodes2008preschoolers; @cimpian2012good; @cimpian2011generic] already suggests that language _can_ have some causal impact in an experimental context. We were interested in whether it actually does, and by what means.

Languages convey gender information in multiple ways including gender-specific proper names, pronouns, and titles (e.g., waiter vs. waitress). In addition, approximately 25% of languages  have some type of grammatical gender [@corbett1991] in which morphological markers are used to signal gender. <!-- The more informative numbers/references may be here.. https://wals.info/chapter/31  It’s more than 25%. Please correct as needed !-->. Some previous work has shown that gender information conveyed by such morphological markers---which are also extended to inanimate objects---can infuse the inanimate objects with “natural” gender [@phillips2003can; @sera1994grammatical]. But more importantly for present purposes, because grammatical gender markers enter into agreement patterns, their use can exaggerate the extent to which gender is being communicated. For example, in Spanish the gender of a nurse has to be signaled grammatically: "enfermer*a*" vs. "enfermer*o*". In Russian, verbs (in some tenses)  are inflected based on the gender of the speaker: "Ya ustal" (I-<span style="font-variant:small-caps;">masc</span> am tired), but "Ya ustala" (I-<span style="font-variant:small-caps;">fem</span> am tired). In both cases, it is not possible to leave the gender unspecified. 

In Study 1a, we examine the degree to which gender in different languages is encoded in their distributional structure. That is, our analysis is not concerned with _what_ is being said about typical gender roles, but rather the statistical patterns to which _potentially_ gendered words are gendered by different languages. In Study 1b we use this information to predict previously collected responses on an gender-career IAT. In Study 2 we examine whether the extent to which different languages use different forms for occupations (e.g., "waiter"/"waitress" but "teacher"/"teacher") correlates with greater implicit gender bias thereby helping to narrow down the source of linguistic knowledge that may be playing a role in shaping gender stereotypes. Together, our data suggest that language not only reflects existing biases, but likely plays a causal role in shaping culturally-specific notions of gender.


```{r child = "sections/study0_writeup.Rmd"}
```

```{r child = "sections/study1_writeup.Rmd"}
```

```{r child = "sections/study2_writeup.Rmd"}
```

# General Discussion
Cultural stereotypes, like gender associations, are pervasive and may have important real-world consequences. Across two studies, we sought to understand the role that language plays in the formation of one particular gender stereotype: the bias to associate men with career and women with family. We estimated the magnitude of this bias among speakers of 25 different languages using a large scale administration of the IAT across over 700,000 people, and then examined how the strength of the stereotype related to gender biases found in language. In Study 1, we found that languages that have a higher degree of gender bias in their co-occurrence statistics tend to have speakers that have stronger gender stereotypes. In Study 2, we found that words that were explicitly marked for the gender of the referent ("waiter"/"waitress") tended to have more biased statistics, and that languages that more frequently  explicitly marked the gender of referents tended to have speakers that were overall more biased. Together these studies provide evidence that the statistics of language use may have a causal influence on the formation of cultural stereotypes.

Our work is the first to characterize the relationship between broad structural patterns in language and cultural stereotypes. Broadly, the positive correlation that we find between these two variables -- gender bias in language and gender bias in speakers -- is consistent with both the possibility that language plays a causal role in the emergence of cultural stereotypes and the idea that language merely reflects existing stereotypes in its speakers. Study 2 however provides suggestive evidence for the causal hypothesis: We find that X. Importantly these two causal forces are not mutually exclusive, and in fact may amplify the effects of each other. For example, the fact that nurses tend to be female in the workforce likely leads speakers to talk about women as nurses more often than men, which in turn produces biased language statistics that may ultimately serve to reinforce a pre-existing cultural stereotype. Future work could use experimental methods to  manipulate language statistics in order to more directly examine these causal influences.

The implications of the language-as-causal-factor hypothesis are important: The mere process of listening to and producing language exposes one to statistics that may lead to the formation of cultural stereotypes. Given that the average adult produces ~16,000 words per day [@mehl2007women], the aggregate effects of exposure to biased language statistics could be large, particularly for children who are just beginning to develop gender stereotypes [@gelman2004mother]. Further, the fact that these language statistics are not explicitly accessible to speakers (versus, for example, the statement, "Most nurses are women.") may make it harder to combat these stereotypes and ultimately exaggerate their effects [@silverstein1981limits]. Future work is needed to understand the relative influence of more explicit messages of gender bias compared to the implicit biases present in language statistics.

An important limitation of the current work is that our measure of the psychological reality of the gender stereotype relies on the IAT, a measure that has been criticized for both its low reliability [@lane2007understanding] and limited validity [e.g., @oswald2013predicting]. In the current work, issues of reliability are less relevant since we are using the IAT to measure group-level differences among large samples of participants. However, issues of validity are an important concern, particularly since we find that languages estimates of bias are uncorrelated with explicit measures. This pattern is in accordance  with the inconsistent relationship between implicit and explicit measures of bias found in the literature [@fazio2003implicit]. In the context of the current work, this limitation could be addressed in future research by using measures of psychological bias that are more closely related to behavior, rather than the IAT.

In conclusion, information about cultural stereotypes is neccessarily acquired through experience in that culture. In our work here, we have suggested that the statistics of language use are an important element of that experience. Many cultural associations and stereotypes are present in the statistics of language may be innocuous -- indeed, these statistics may be an important mechanism through which cultural information is transmitting. But, in other cases, like the kind of gender stereotypes investigated here, language may be playing a powerful role in their formation, and ultimately contributing to undesirable real-world consequences like gender inequality in STEM. Understanding  the causal role that language plays in the formation of these stereotypes is therefore an important first step to changing these consequences.
 
```{r}
# IAT. Nevertheless, our work is an important first step in demonstrating a link between structural properties of language and cultural stereotypes across a large sample of speakers and languages.
 
```

\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

---
nocite: | 
  @caliskan2017semantics
  @ciafactbook
  @nosek
  @stoet2018gender
  @miller2015women
...


\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
