---
title: Data readme and results for Study 1c 
author: Molly Lewis 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: no
    toc: no
---
  
******

```{r setup, include = F}

# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(here)
library(broom)
library(langcog)
library(lme4)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)
```

We have estimates of the language IAT bias for each IAT (N = 31) from 5 different training runs of each of the two corpora (BNC and COCA). 
```{r}
LANGUAGE_PATH <- here("data/study1c/processed/bnc_vs_coca_es_400_10_x5.csv")

# language es (5 runs of each model)
es_lang_raw <- read_csv(LANGUAGE_PATH) 
es_lang_tidy <- es_lang_raw %>%
  spread(model, effect_size) %>%
  rename(coca_lang_es = coca, 
         bnc_lang_es = bnc) %>%
  mutate(lang_diff = bnc_lang_es - coca_lang_es) # get bnc - coca language es difference %

```



### Behavioral effect size
FILENAME: `tidy_behavioral_iat_data.csv`

We have estimates of behavioral bias for each IAT for each sample of participants (UK and USA).
```{r}
RAW_BEHAVIORAL_CONF <- here("data/study1c/processed/tidy_behavioral_iat_data_confirmatory_full.csv") 

behavioral <- read_csv(RAW_BEHAVIORAL_CONF) %>%
  select(user_id, domain, residence, resid)%>%
  rename(behavioral_effect_resid = resid) %>%
  mutate(user_id = as.character(user_id)) 

full_df <- behavioral %>%
  mutate(temp = list(es_lang_tidy)) %>%
  unnest() %>%
  filter(domain == domain1) %>%
  select(-domain1) %>%
  mutate_if(is.character, as.factor) %>%
  mutate_if(is.numeric, scale) 

#write_csv(full_df, "long_form_confirmatory_data.csv")
```

```{r}
behavioral_effect_resid.lmer <- full_df %>% 
    group_by(user_id, domain,residence) %>%
    mutate(residence = fct_rev(residence)) %>%
    summarize_if(is.numeric,mean) %>%
    lmer(behavioral_effect_resid ~ residence*lang_diff+(1|user_id)+(1|domain),data=.)

summary(behavioral_effect_resid.lmer)
```

```{r, misc}
lm(behavioral_effect_resid ~ residence:lang_diff, data = full_df) %>%
  summary()

 # mutate(language_es = ifelse(residence == "us", coca_lang_es,
 #                             bnc_lang_es))

model1 <- lmer(behavioral_effect_resid ~ residence*lang_diff +  (1|run) + (1|domain), data = full_df)
summary(model1)

model2 <- lmer(behavioral_effect_resid ~ language_es + residence + (1|run) + (1|domain), data = full_df)
summary(model2)




long_lang <- es_lang_tidy %>% 
  select(-lang_diff)  %>%
  rename(us = coca_lang_es, 
         uk = bnc_lang_es) %>%
  gather("language", "language_bias", -run, -domain)
  
full_df <- behavioral %>%
  left_join(long_lang, by = c("residence" = "language")) %>%
  mutate(residence = ifelse(residence == "uk", -0.5, 0.5))

model <- lmer(behavioral_effect_resid ~ lang_diff*residence +   (1|run) + (1|domain) + (1|user_id), data = full_df)

lmer(behavioral_effect_resid ~ language_bias:domain +  (1|run), data = full_df ) %>%
  summary()


lmer(D ~ language_bias:domain + sex+education+age+block_order+ (1|run), data = full_df ) %>%
  summary()
```

```{r}


full_df <- behavioral %>%
  mutate(temp = list(es_lang_tidy)) %>%
  unnest() %>%
  mutate_if(is.character, as.factor) %>%
  mutate_if(is.numeric, scale)


lmer(behavioral_effect_resid ~ bnc_lang_es*residence + coca_lang_es*residence +  (1|domain) + (1|run), data = full_df ) %>%
  summary()

lmer(D ~ bnc_lang_es*residence + coca_lang_es*residence + sex+education+age+block_order+ (1|domain) + (1|run), data = full_df ) %>%
  summary()
