---
title: Montag children's book corpus
subtitle: training w2v on corpus
author: Molly Lewis 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: no
    toc: yes
---
  
******

```{r setup, include = F}

# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(tidytext)
library(langcog)
library(data.table)
library(wordVectors)
library(corpus)
library(kableExtra)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)
```
FINDINGS:

- The N closest words to "boy" and "girl" are almost non-overlapping, suggesting that the concepts of maleness and femaleness are located in different parts of semantic space (maybe related to valence?)
- There are more male-biased books than female-biased books (61% vs. 39%).
- There's less variability in semantic space for female-biased books, compared to male-biased books.


## Train word2vec model on montag corpus.
```{r, eval = F, include = F}
d_raw <- read_lines("100Books.txt")  %>%
  as.data.frame() %>%
  rename("text" = ".")  %>%
      filter(!str_detect(text, regex("^Title",
                                                 ignore_case = TRUE)),
           !str_detect(text, regex("^Author",
                                                 ignore_case = TRUE)),
           text != "")

PATH <- "100Books_clean.txt"
#write_delim(d_raw, PATH)

prep_word2vec(origin = PATH,destination="books_prepped.txt",lowercase=T,bundle_ngrams=1)
```

```{r  include = F}
model <- train_word2vec("books_prepped.txt",
                        "childers_book_vectors_uni.bin", 
                        vectors=200,threads=4,
                        window=12,iter=5,
                        negative_samples=0, 
                        min_count = 5, force = T)
model_df <- model %>%
  as.data.frame() %>%
  rownames_to_column("target_word")

#write_csv(model_df, "w2v_montag_corpus.csv")
```

### Book centroids
This is the distribution of books in semantic space using t-sne coordinates. The cenroid of each book is calculated as the sum of all words in book.
```{r}
model_df <- read_csv("w2v_montag_corpus.csv")

montag_counts <- read_csv("tidy_montag_corpus.csv") %>%
  rename(target_word = word) %>%
  count(title, target_word)

# get location of essay in the vector space 
get_one_essay_location <- function(book_title, 
                                   model, 
                                   book_data){
  
  # merge current essay with corresponding word vecs
  merged_vectors <- book_data[title == book_title] %>%
    merge(model, all.x = TRUE, by = "target_word")
  
  # get pieces to essay vector computation
  word_vectors <- merged_vectors %>%  # get word vectors only
    select(V1:V200) %>%
    as.matrix()
  counts <- merged_vectors$n # get counts

  # weighted by number of times word appears in essay
  count_weighted_vecs <- word_vectors * counts # weight by n word appears
  essay_vector_count_weighted <- t(colSums(count_weighted_vecs, 
                                           na.rm = TRUE)) # sum across words
 
  # return essay vector
  count_weighting_vec <- data.frame(title = book_title,
                                    essay_vector_count_weighted)

}

essay_locations <- 
  map_df(unique(montag_counts$title), 
         get_one_essay_location, model_df, 
         as.data.table(montag_counts))

# get tsne coordinates
tsne_out = Rtsne::Rtsne(as.matrix(essay_locations[,-1]))
tsne_dims <- tsne_out$Y %>%
  as.data.frame() %>%
  rename(tsne_X = V1,
         tsne_Y = V2)  

#write_csv(tsne_dims, "book_tsne_montag_w2v.csv")
```

```{r}
tsne_dims <- read_csv("book_tsne_montag_w2v.csv") %>%
  bind_cols(title = unique(montag_counts$title)) 

ggplot(tsne_dims,
         aes(x = tsne_X, y = tsne_Y)) +
  geom_text(aes(label = title), size = 2) +
  theme_void()
```

## Relative position of gender-related words
Get N closest words (N = 200) to "boy" and "girl" word pairs. Then, plot in 2d space and try to predict words with independent ratings (Glasglow and Byrsbaert).

```{r}
# Read in w2v and norms.
gender_norms <- read_csv("GlasgowNorms.csv") %>%
  select(word, GEND_M) %>%
  rename(gender_maleness = GEND_M)

affect_norms <- read_csv("BRM-emot-submit.csv") %>%
  select(2,3,6,9) %>%
  rename(word = Word,
         valence = V.Mean.Sum, 
         affect = A.Mean.Sum,
         dominance = D.Mean.Sum)
```

### Boy vs. girl
#### Distribution in 2d w2v space

The large point corresponds to the target boy and girl words.

```{r, fig.height = 6}
BOY_WORD <- "boy"
GIRL_WORD <- "girl"

boy_words <- model %>% 
  closest_to(BOY_WORD, n = 201, fancy_names = F) %>%
  slice(-1) %>%
  mutate(word_type = "boy") 
  #mutate(word = text_tokens(word, stemmer = "en") %>% unlist())

girl_words <- model %>% 
  closest_to(GIRL_WORD, n = 201, fancy_names = F) %>%
  slice(-1) %>%
  mutate(word_type = "girl") 
  #mutate(word = text_tokens(word, stemmer = "en") %>% unlist())

word_dims <- boy_words %>%
  bind_rows(girl_words) %>%
  select(-similarity, -word_type) %>%
  add_row(word = GIRL_WORD) %>%
  add_row(word = BOY_WORD) %>%
  rename(target_word = word) %>%
  left_join(model_df) %>%
  distinct(target_word, .keep_all=T)

# get tsne coordinates
tsne_out = Rtsne::Rtsne(as.matrix(word_dims[,-1]))
tsne_dims <- tsne_out$Y %>%
  as.data.frame() %>%
  rename(tsne_X = V1,
         tsne_Y = V2)  %>%
  bind_cols(target_word = word_dims$target_word) %>%
  select(target_word, everything()) 

tsne_words <- boy_words %>%
  bind_rows(girl_words) %>%
  select(-similarity) %>% 
  rename(target_word = word) %>%
  left_join(tsne_dims)

targ_words <- tsne_dims %>%
  filter(target_word %in% c(BOY_WORD, GIRL_WORD)) %>%
  mutate(word_type = ifelse(target_word == BOY_WORD,
                            "boy", "girl"))
 
ggplot(tsne_words,
         aes(x = tsne_X, y = tsne_Y)) +
  geom_text(aes(label = target_word, color = word_type),
            size = 2) +
  scale_color_manual(values = c("blue", "red")) +
  geom_point(aes(color = word_type), 
             data = targ_words, size = 4) +
  theme_void()
```

#### Independent ratings
```{r, fig.height = 3}
boy_words %>%
  bind_rows(girl_words) %>%
  left_join(gender_norms) %>%
  left_join(affect_norms) %>%
  gather("norm", "rating", -1:-3) %>%
  group_by(norm, word_type) %>%
  multi_boot_standard(col = "rating", na.rm = T) %>%
  ggplot(aes(x = fct_rev(word_type), y = mean, 
             fill = word_type)) +
    facet_wrap(~norm, nrow = 1) +
    geom_bar(stat = "identity") +
    ggtitle(paste0(BOY_WORD, " vs. ", GIRL_WORD)) +
    ylab("Mean rating") +
    xlab("gender of target words") +
    ylim(0, 7) +
    geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
    theme_minimal() +
    theme(legend.position = "none")
```


### Him vs. her
#### Distribution in 2d w2v space
```{r, fig.height = 6}
BOY_WORD <- "him"
GIRL_WORD <- "her"

boy_words <- model %>% 
  closest_to(BOY_WORD, n = 201, fancy_names = F) %>%
  slice(-1) %>%
  mutate(word_type = "boy") 
  #mutate(word = text_tokens(word, stemmer = "en") %>% unlist())

girl_words <- model %>% 
  closest_to(GIRL_WORD, n = 201, fancy_names = F) %>%
  slice(-1) %>%
  mutate(word_type = "girl") 
  #mutate(word = text_tokens(word, stemmer = "en") %>% unlist())

word_dims <- boy_words %>%
  bind_rows(girl_words) %>%
  select(-similarity, -word_type) %>%
  add_row(word = GIRL_WORD) %>%
  add_row(word = BOY_WORD) %>%
  rename(target_word = word) %>%
  left_join(model_df) %>%
  distinct(target_word, .keep_all=T)

# get tsne coordinates
tsne_out = Rtsne::Rtsne(as.matrix(word_dims[,-1]))
tsne_dims <- tsne_out$Y %>%
  as.data.frame() %>%
  rename(tsne_X = V1,
         tsne_Y = V2)  %>%
  bind_cols(target_word = word_dims$target_word) %>%
  select(target_word, everything()) 

tsne_words <- boy_words %>%
  bind_rows(girl_words) %>%
  select(-similarity) %>% 
  rename(target_word = word) %>%
  left_join(tsne_dims)

targ_words <- tsne_dims %>%
  filter(target_word %in% c(BOY_WORD, GIRL_WORD)) %>%
  mutate(word_type = ifelse(target_word == BOY_WORD,
                            "boy", "girl"))
 
ggplot(tsne_words,
         aes(x = tsne_X, y = tsne_Y)) +
  geom_text(aes(label = target_word, color = word_type),
            size = 2) +
  scale_color_manual(values = c("blue", "red")) +
  geom_point(aes(color = word_type), 
             data = targ_words, size = 4) +
  theme_void()
```

#### Independent ratings
```{r, fig.height = 3}
boy_words %>%
  bind_rows(girl_words) %>%
  left_join(gender_norms) %>%
  left_join(affect_norms) %>%
  gather("norm", "rating", -1:-3) %>%
  group_by(norm, word_type) %>%
  multi_boot_standard(col = "rating", na.rm = T) %>%
  ggplot(aes(x = fct_rev(word_type), y = mean, 
             fill = word_type)) +
    facet_wrap(~norm, nrow = 1) +
    geom_bar(stat = "identity") +
    ggtitle(paste0(BOY_WORD, " vs. ", GIRL_WORD)) +
    ylab("Mean rating") +
    ylim(0, 7) +
    xlab("gender of target words") +

    geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
    theme_minimal() +
    theme(legend.position = "none")
```

### His vs. her
#### Distribution in 2d w2v space
```{r, fig.height = 6}
BOY_WORD <- "his"
GIRL_WORD <- "her"

boy_words <- model %>% 
  closest_to(BOY_WORD, n = 201, fancy_names = F) %>%
  slice(-1) %>%
  mutate(word_type = "boy") 
  #mutate(word = text_tokens(word, stemmer = "en") %>% unlist())

girl_words <- model %>% 
  closest_to(GIRL_WORD, n = 201, fancy_names = F) %>%
  slice(-1) %>%
  mutate(word_type = "girl") 
  #mutate(word = text_tokens(word, stemmer = "en") %>% unlist())

word_dims <- boy_words %>%
  bind_rows(girl_words) %>%
  select(-similarity, -word_type) %>%
  add_row(word = GIRL_WORD) %>%
  add_row(word = BOY_WORD) %>%
  rename(target_word = word) %>%
  left_join(model_df) %>%
  distinct(target_word, .keep_all=T)

# get tsne coordinates
tsne_out = Rtsne::Rtsne(as.matrix(word_dims[,-1]))
tsne_dims <- tsne_out$Y %>%
  as.data.frame() %>%
  rename(tsne_X = V1,
         tsne_Y = V2)  %>%
  bind_cols(target_word = word_dims$target_word) %>%
  select(target_word, everything()) 

tsne_words <- boy_words %>%
  bind_rows(girl_words) %>%
  select(-similarity) %>% 
  rename(target_word = word) %>%
  left_join(tsne_dims)

targ_words <- tsne_dims %>%
  filter(target_word %in% c(BOY_WORD, GIRL_WORD)) %>%
  mutate(word_type = ifelse(target_word == BOY_WORD,
                            "boy", "girl"))
 
ggplot(tsne_words,
         aes(x = tsne_X, y = tsne_Y)) +
  geom_text(aes(label = target_word, color = word_type),
            size = 2) +
  scale_color_manual(values = c("blue", "red")) +
  geom_point(aes(color = word_type), 
             data = targ_words, size = 4) +
  theme_void()
```

#### Independent ratings
```{r, fig.height = 3}
boy_words %>%
  bind_rows(girl_words) %>%
  left_join(gender_norms) %>%
  left_join(affect_norms) %>%
  gather("norm", "rating", -1:-3) %>%
  group_by(norm, word_type) %>%
  multi_boot_standard(col = "rating", na.rm = T) %>%
  ggplot(aes(x = fct_rev(word_type), y = mean, 
             fill = word_type)) +
    facet_wrap(~norm, nrow = 1) +
    geom_bar(stat = "identity") +
    ggtitle(paste0(BOY_WORD, " vs. ", GIRL_WORD)) +
    ylab("Mean rating") +
    xlab("gender of target words") +
    ylim(0, 7) +
    geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
    theme_minimal() +
    theme(legend.position = "none")
```

## Gender score of each word
To calculate gender bias score for each word,  I took the mean distance to 4 “male” words (“boy”, “man”,  “he”, “him”) and the mean distance to 4 “female” words (“girl”, “woman”,  “she”, “her”), and subtracted the two. Positive numbers indicate male bias. 

```{r}
w2v_norms <- read_csv("gender_score_output_montag_trained.csv", 
                      col_names = c("target_word", "female_score",
                                    "male_score", "gender_score")) %>%
  distinct(.keep_all = T) %>%
  anti_join(stop_words %>% filter(lexicon == "SMART"), 
            by = c("target_word" = "word"))

ggplot(w2v_norms, aes(x = gender_score)) +
    ggtitle("Gender score distribution from word embedding models")+
  xlab("Gender score + => more male") +
    geom_histogram() +
    theme_classic()
```

Glasglow norms:
```{r}
norms <- read_csv("GlasgowNorms.csv") %>%
  select(word, GEND_M, contains("_M")) 

all_norms <- left_join(w2v_norms, norms, by = c("target_word"= "word")) %>%
  rename(gender_rating = GEND_M) %>%
  select(target_word, gender_score, gender_rating)


ggplot(all_norms,aes(y = gender_score, x = gender_rating)) +
    geom_point() +
    geom_smooth(method = "lm")  +
    ggtitle("Correlation between gender norms and word embedding gender score")+
    ylab("word2vec gender score") +
    scale_x_continuous(name = "mean rating", 
                     limits = c(1, 7),
                     breaks = 1:7) +
  theme_classic()

cor_summary <- cor.test(all_norms$gender_score, all_norms$gender_rating)
cor_summary
```

These are correlated with the Glasglow norms, though not as strongly as when using the Wikipedia-trained corpus.

## Gender score by book

Here I calculated a gender score for each book as the mean gender score of all the words in the book (unweighted for word frequency.)
```{r}
montag <- read_csv("tidy_montag_corpus.csv") %>%
  rename(target_word = word)

tidy_with_norms <- montag %>%
  left_join(w2v_norms)

norm_means <- tidy_with_norms %>%
  group_by(title) %>%
  multi_boot_standard(col = "gender_score", na.rm = T)

norm_means_wide <- norm_means %>%
  select(title, mean) %>%
  arrange(mean)

kable(norm_means_wide, "html") %>%
  kable_styling() %>%
  scroll_box(width = "900px", height = "400px")
```

```{r}
ggplot(norm_means_wide, aes(x = mean)) +
    geom_histogram() +
    ggtitle("mean rating by book") +
    xlab("word2vec gender rating (+ => more male)") +
    geom_vline(aes(xintercept = 0), color = "red", linetype = 2) +
    theme_classic()

#norm_means_wide %>%
 # mutate(male = ifelse(mean > 0, 1, 0)) %>%
 # ungroup()%>%
 # count(male)
  
```


```{r}
tsne_dims <- read_csv("book_tsne_montag_w2v.csv") %>%
  bind_cols(gender_rating = norm_means$mean)  %>%
  bind_cols(title = norm_means$title)  %>%
  mutate(gender_rating_cat = ifelse(gender_rating > 0, 
                                    "male", "female"))


  ggplot(tsne_dims,
         aes(x = tsne_X, y = tsne_Y, color = gender_rating)) +
  geom_text(aes(label = title, color = gender_rating), size = 2) +
  #geom_point() +
  scale_color_gradient2(low = "red", high = "blue") +
    theme_void()
  
  
  ggplot(tsne_dims,
         aes(x = tsne_X, y = tsne_Y, color = gender_rating)) +
    geom_point() +
  #geom_text(aes(label = title, color = gender_rating), size = 2) +
  #geom_point() +
  scale_color_gradient2(low = "red", high = "blue") +
    theme_void()
```
