---
title: Study 1C analyses
subtitle: ED8
date: "`r Sys.Date()`"
output: latex_document
---


```{r setup, include = F}
# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(here)
library(broom)
opts_chunk$set(echo = F, message = F, warning = F,
               error = F, tidy = F, cache = F)
```

`


```{r}
LANG_MEASURE_PATH <- here("writeup/journal/supporting_information/ED/ED8/ED8_data.csv")
```



## Predicting career-gender association with both language measures 
In this analysis, we predict the magnitude of implicit career-gender association by language with an additive linear model. As predictors, we include   proportion gender distinct labels and linguistic career-gender association (as measured by word embeddings of the IAT words). Model coefficients are shown below for models based on the Subtitle (top) and Wikipedia (bottom) corpora. 

<b> Subtitle Corpus: </b>
```{r}
all_lang_model_scaled <- read_csv(LANG_MEASURE_PATH)

# regressions with lang_es_sub/wiki instead of occupation statistics
subt_regression_r2_no_age <- lm(es_iat_sex_age_order_implicit_resid ~ mean_prop_distinct_occs + 
     lang_es_sub, 
   data = all_lang_model_scaled)

subt_regression_r2_no_age  %>%
  tidy() %>%
  mutate(term = fct_recode(term,
         `Lang. Male-Career Assoc (Subtitle)` = "lang_es_sub",
         `Prop. Gendered Occupuation Terms` = "mean_prop_distinct_occs")) %>%
  kable(digits = 3) 
```
This model accounts for `r round(summary(subt_regression_r2_no_age)$r.squared * 100, 2)`% variance. 

<b> Wikipedia Corpus: </b>
```{r}
wiki_regression_r2_no_age <- lm(es_iat_sex_age_order_implicit_resid ~ mean_prop_distinct_occs + 
     lang_es_wiki, 
   data = all_lang_model_scaled)  

wiki_regression_r2_no_age %>%
  tidy() %>%
  mutate(term = fct_recode(term,
         `Lang. Male-Career Assoc (Wikipedia)` = "lang_es_wiki",
         `Prop. Gendered Occupuation Terms` = "mean_prop_distinct_occs")) %>%
  kable(digits = 3) 
```
This model accounts for `r round(summary(wiki_regression_r2_no_age)$r.squared * 100, 2)`% variance. 


Presented below are models that include median country age as an additional predictor. 


## Title and Legend 
Models predicting career-gender association with proportion gender distinct labels and linguistic career-gender association (Study 2). In this analysis, we predict the magnitude of implicit career-gender association by language with an additive linear model. As predictors, we include   proportion gender distinct labels and linguistic career-gender association (as measured by word embeddings of the IAT words). Model coefficients are shown below for models based on the Wikipedia (a) and Subtitle (b) corpora. 
