---
title: GDP control analyses
author: Molly Lewis and Gary Lupyan
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: no
    toc: yes
---
  
******

```{r setup, include = F}

# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(here)
library(ggrepel)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)

```

## Correlations between all measures at the country level

`gdp_2017` is per capita GDP taken from the World Bank database (the 2017 "GDP per capita (current US$)" indicator); it is missing values for one country (Taiwan). This measure is correlated with the GDP measure from 2013 (`gdp_2013`) at _r_ = .96. Models below are presented with both the 2013 and 2017 measures. Note that all of the variables are centered and scaled.

```{r}
# get GDP 2017 data from World Bank API
gdp_data <- wbstats::wb(indicator = "NY.GDP.PCAP.CD", 
                        startdate = 2017, 
                        enddate = 2017) %>%
            select(iso2c, value) %>%
            rename(gdp_2017 = value)

pref_data <- readstata13::read.dta13("genderdifferences.dta") %>%
  mutate(country_code = countrycode::countrycode(ison, "iso3n", "iso2c")) %>%
  select(country_code, everything()) %>%
  select(-ison)

STOET_PATH <- "/Users/mollylewis/Documents/research/Projects/1_in_progress/IATLANG/exploratory_studies/7_age_controls/stoet_data.csv"
stoet_data <- read_csv(STOET_PATH) %>%
  mutate(country_code = countrycode::countrycode(country_name, "country.name", "iso2c")) %>%
  select(country_code, everything())

# data from Bill von Hippel
INPATH <- "data/Molly data2.csv" 
country_raw <- read_csv(INPATH) %>%
  janitor::clean_names()  %>%
  left_join(gdp_data, by = c("country_code" = "iso2c"))   %>%
  left_join(stoet_data) %>%
  left_join(pref_data)

# save country data with GDP 2017 data merged in (unscaled)
# OUTPATH <- "country_level_data_with_GDP.csv"
# write_csv(country_raw, OUTPATH)

# scale variables
country_level <- country_raw %>%
  mutate_if(is.numeric, base::scale) 
```

```{r corrplot, fig.height = 10, fig.width = 10, fig.cap = "Pairwise correlation between all country-level measures. Red and blue correspond to positive and negative correlations, respectively. Non-significant correlations (\\textit{p} >= .05) are indicated with white squares."}
plot_data <- country_level %>%
  select_if(is.numeric) %>%
  select(-n_participants, -gdp_2013)

long_corr <- cor(plot_data, 
                use = "pairwise.complete.obs") %>%
  as.data.frame() %>%
  rownames_to_column("v2") %>%
  gather("v1", "estimate", -v2)

long_p <- corrplot::cor.mtest(plot_data, 
                             use = "pairwise.complete.obs")$p %>%
  as.data.frame(row.names = names(plot_data)) %>%
  do(setNames(.,names(plot_data))) %>%
      rownames_to_column("v2") %>%
  gather("v1", "p", -v2)

corr_df <- full_join(long_corr, long_p) %>%
  mutate(estimate_char = case_when(v1 == v2 ~ "", 
                              TRUE ~ as.character(round(estimate,2))),
         estimate = case_when(v1 == v2 ~ as.numeric(NA), 
                              TRUE ~ estimate),
         estimate_color = case_when(p < .05 ~ estimate, TRUE ~ 0),
         v1 = fct_relevel(v1, "lang_es_sub", "lang_es_wiki", "subt_occu_semantics_fm",
                          "wiki_occu_semantics_fm", "mean_prop_distinct_occs", "implicit_resid", "explicit_resid", "median_country_age", "gdp_2017", "per_women_stem", "gender_inequality_index_value", "science_literacy_diff", "intra_indv_diff", "self_efficacy_diff", "intrest_diff", "enjoy_diff", "satisfaction"),
           v2 = fct_relevel(v2, "lang_es_sub", "lang_es_wiki", "subt_occu_semantics_fm",
                          "wiki_occu_semantics_fm", "mean_prop_distinct_occs", "implicit_resid", "explicit_resid", "median_country_age", "gdp_2017", "per_women_stem", "gender_inequality_index_value", "science_literacy_diff", "intra_indv_diff", "self_efficacy_diff", "intrest_diff", "enjoy_diff","satisfaction"))

ggplot(corr_df, aes(v1, fct_rev(v2), fill = estimate_color)) + 
  geom_tile() + #rectangles for each correlation
  #add actual correlation value in the rectangle
  geom_text(aes(label = estimate_char), size = 3) + 
  scale_fill_gradient2(low ="blue", mid = "white", high = "red", 
                       midpoint = 0, space = "Lab", guide = "colourbar",
                       name = "Pearson's r") +
  ggtitle("Pairwise Correlation Coefficients") +
  theme_classic(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), #, hjust = .95, vjust = .2), 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")
```

## Target correlations{.tabset}

### STEM ~ language bias
```{r, fig.width = 5}
country_level %>%
  ggplot(aes(x = lang_es_sub, y = women_stem, label = country))+
  geom_point() +
  geom_text_repel(size = 3) +
  ylab("Perc. Women in STEM") +
  xlab("Linguistic Gender Bias\n(effect size)") +
  ggtitle("Perc. Women in STEM vs. Language Bias") +
  geom_smooth(method = "lm", alpha = .2) +
  theme_classic()
```

### GDP ~ language bias
```{r, fig.width = 5}
country_level %>%
  ggplot(aes(x = lang_es_sub, y = gdp_2017, label = country))+
  ylab("GDP (2017)") +
  xlab("Linguistic Gender Bias\n(effect size)") +
  ggtitle("GDP vs. Language Bias") +
  geom_smooth(method = "lm", alpha = .2) +
  geom_point() +
  geom_text_repel(size = 3) +
  theme_classic(base_size = 12)
```

### GDP ~ STEM
```{r, fig.width = 5}
country_level %>%
  ggplot(aes(x = women_stem, y = gdp_2017, label = country))+
  ylab("GDP (2017)") +
  xlab("Perc. Women in STEM") +
  ggtitle("GDP vs. Perc. Women in STEM ") +
  geom_smooth(method = "lm", alpha = .2) +
  geom_point() +
  geom_text_repel(size = 3) +
  theme_classic(base_size = 12)
```


## Models predicting gender-career IAT from language biases{.tabset}

### GDP 2017
```{r}
lm(implicit_resid ~ lang_es_sub + gdp_2017,  
   data = country_level) %>%
  summary()
```


### GDP 2013
```{r}
lm(implicit_resid ~ lang_es_sub + gdp_2013,  
   data = country_level) %>%
  summary()
```


## Models predicting women-in-STEM bias from language biases{.tabset}

### GDP 2017
```{r}
lm(women_stem ~  lang_es_sub + gdp_2017,  
   data = country_level) %>%
  summary()
```


### GDP 2013
```{r}
lm(women_stem ~  lang_es_sub + gdp_2013,  
   data = country_level) %>%
  summary()
```

```{r, eval = F, include = F}

df <- read_csv(OUTPATH) %>%
  filter(!is.na(gdp_2017), !is.na(lang_es_sub), !is.na(women_stem))

lm(women_stem ~  lang_es_sub + gdp_2017,  
   data = df) %>%
  summary()

participant_level <- read_csv(here("data/study0/processed/by_participant_df.csv")) %>%
  select(country_code, es_iat_sex_age_order_implicit_resid, sex, log_age, order, overall_iat_D_score)

participant_level_with_country_lang <- participant_level %>%
  left_join(country_level %>% select(country_code, gdp_2017, language_name)) %>%
  left_join(lang_level %>% select(language_name, lang_es_sub)) 

lmer(es_iat_sex_age_order_implicit_resid ~  lang_es_sub + gdp_2017 + (1|country_code)+(1|language_name), data=participant_level_with_country_lang) %>%
  summary()

lmer(overall_iat_D_score ~  lang_es_sub + gdp_2017 + log_age + order + sex + (1|country_code)+(1|language_name), data=participant_level_with_country_lang) %>%
  summary()
  

```

```{r}
lm(ggi ~ self_efficacy_diff + lang_es_sub   ,  
   data = country_level) %>%
  summary()

lm(ggi ~   self_efficacy_diff   ,  
   data = country_level) %>%
  summary()

lm(ggi ~   lang_es_sub   ,  
   data = country_level) %>%
  summary()
library(robmed)

mediation_model_bootsrapped <- country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "ggi",
    m = "self_efficacy_diff",
    covariates = c("gdp_2017", "median_country_age")) %>%
   test_mediation()

mediation_model_bootsrapped
p_value(mediation_model_bootsrapped)
```

```{r}

#**
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "per_women_stem",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()

#**
country_level %>%
  fit_mediation(
    x = "lang_es_wiki",
    y = "per_women_stem",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()

# **
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "ggi",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()

country_level %>%
  fit_mediation(
    x = "lang_es_wiki",
    y = "ggi",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()

# **
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "gini",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()

country_level %>%
  fit_mediation(
    x = "lang_es_wiki",
    y = "gini",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()

# **
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "hdi_value",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()

country_level %>%
  fit_mediation(
    x = "lang_es_wiki",
    y = "hdi_value",
    m = "self_efficacy_diff") %>%
   test_mediation() %>%
   p_value()



# +
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "per_women_stem",
    m = "implicit_resid",
     covariates = c("median_country_age", "gdp_2017")) %>%
   test_mediation() %>%
   p_value()


# +
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "per_women_stem",
    m = "explicit_resid",
    covariates = c("median_country_age", "gdp_2017")) %>%
   test_mediation() %>%
   p_value()

```

lang -> implicit -> self_efficacy
```{r}
# +
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "self_efficacy_diff",
    m = "implicit_resid") %>%
   test_mediation() %>%
   p_value()

# +
country_level %>%
  fit_mediation(
    x = "lang_es_wiki",
    y = "self_efficacy_diff",
    m = "implicit_resid",
    covariates = c("median_country_age", "gdp_2017")) %>%
   test_mediation() %>%
   p_value()

```


```{r}

#**
country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "gdp_2017",
    m = "genderdif") %>%
   test_mediation() %>%
   p_value()

country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "per_women_stem",
    m = "genderdif") %>%
   test_mediation() %>%
   p_value()

country_level %>%
  fit_mediation(
    x = "lang_es_sub",
    y = "ggi",
    m = "genderdif") %>%
   test_mediation() %>%
   p_value()

country_level %>%
  fit_mediation(
    x = "lang_es_wiki",
    y = "ggi",
    m = "genderdif") %>%
   test_mediation() %>%
   p_value()



country_level %>%
  fit_mediation(
    x = "genderdif",
    y = "ggi",
    m = "lang_es_wiki") %>%
   test_mediation() %>%
   p_value()
```

