---
title             : "What are we learning from language? Associations between gender biases and distributional semantics in 25 languages"
shorttitle        : "What are we learning from language?"
header-includes: 
- \usepackage{float} 
- \floatplacement{figure}{t!} # make every figure with caption = t

author: 
  - name          : "Molly Lewis"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "mollyllewis@gmail.com"
    address       : ""
  - name          : "Gary Lupyan"
    affiliation   : "1"
    
affiliation: 
  - id            : "1"
    institution   : "University of Wisconsin-Madison"
  - id            : "2"
    institution   : "University of Chicago"

authornote: |
    Portions of this manuscript appeared in Lewis & Lupyan, 2018, Cog. Sci. Proceedings. 

abstract: | 
  Cultural stereotypes such as the idea that men are more suited for paid work while women for taking care of the home and family may contribute to gender imbalances in STEM fields (e.g., Leslie, Cimpian, Meyer, & Freeland, 2015) and other undesirable gender disparities. Here, we test the hypothesis that word co-occurrence statistics (e.g., the co-occurrence of “nurse” with “she”) play a causal role in the formation of the men-career/women-family stereotype. We use word embedding models to measure bias in the distributional statistics of 25 languages and find that languages with larger biases tend to have speakers with larger implicit biases (_N_ = 657,335). These biases are further related to the extent that languages mark gender in their lexical forms (e.g., “waiter”/“waitress”) hinting that linguistic biases may be causally related to biases shown in people's implicit judgments.
  
keywords          : "cultural stereotypes, implicit association task (IAT), gender"
wordcount         : "1998 (excluding methods/results)"

bibliography      : ["library.bib"]

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library(papaja)
library(kableExtra)
library(here)
library(numform)
source(here("writeup/journal/helpers/psych_to_mat.R"))
```

```{r analysis-preferences, include = F}
set.seed(42) # Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, 
                      warning = FALSE, 
                      message = FALSE, 
                      cache = FALSE,
                      echo = FALSE,
                      fig.pos = "t!")
```

# Introduction
By the time children are two years old, they have begun to acquire the gender stereotypes in their culture [Gelman et al., -@gelman2004mother]. These stereotypes can have undesirable effects. For example, in one study, 6-year-old girls were less likely than boys to choose activities that were described as for children “who are very, very smart” and also less likely to think of themselves as “brilliant” [@bian2017gender]. Such beliefs may, over time, translate to the observed lower rates of female participation in STEM fields [@miller2015women; @stoet2018gender; @ceci2011understanding; @leslie2015expectations].  For this reason and others, it is important to better understand how cultural stereotypes are formed. 

We can distinguish between two major sources of information on which gender stereotypes may be based. The first is direct experience. For example, one may observe that most nurses are women and most philosophers are men and conclude that women are better suited for nursing and men for philosophy. The second is language. Even without any direct experience with nurses or philosophers, one may learn about their stereotypical gender from language about nurses and philosophers. Languages encode gender in multiple ways. These include gender-specific titles ("Mr". vs. "Miss"), proper names (“Sam” vs. “Ashley”), pronouns (“he” vs. “she”), certain job titles (“waiter” vs. “waitress”), and higher-order linguistic associations (otherwise gender-neutral words can become gendered by being associated with explicitly gendered contexts). Another source of linguistic information comes from sex-based grammatical gender systems found in approximately 30% of languages (Dryer \& Haspelmath, 2013). For example, in Spanish, the gender of a nurse must be specified grammatically (“enfermer*a*” vs. “enfermer*o*”).

To the extent that language is a source of information for forming cultural stereotypes, two people with similar direct experiences, but different linguistic experiences, may develop different stereotypes. Some past work hints at people’s surprising sensitivity to stereotype-relevant information delivered through language. Young children perform worse in a game if they are told that someone of the opposite gender performed better than they did on a previous round [@rhodes2008preschoolers], or merely told that the game is associated with a particular gender  [@cimpian2012good]. In some cases, a subtle turn of phrase can influence children’s gender-based generalization [@rhodes2018subtle; @cimpian2011generic]. For example, Cimpian and Markman found that children were more likely to infer that a novel skill is stereotypical of a gender if the skill is introduced with a generic as opposed a non-generic subject (“[Girls are/There is a girl who is] really good at a game called ‘gorp”’). Such work shows that in certain experimental settings, language can influence stereotype formation. We were interested in whether it actually does, and by what means. 

A widely used method for quantifying cultural stereotypes at an individual level is the _Implicit Association Test_ [IAT; @greenwald1998measuring]. Here, we use previously administered IATs  designed to measure a particular type of gender stereotype: A bias to associate men with careers and women with family (_N_ = 657,335; Nosek, et al., 2002). These data span native speakers of 25 languages allowing us to assess how performance varies with properties of languages.

Discovering that linguistic bias predicts people’s implicit biases can be interpreted in at least two ways. The first is that some cultures have stronger stereotypes and these are reflected in what people talk about. Language, on this view, simply _reflects_ pre-existing biases. We refer to this as the _language as reflection_ hypothesis. However, language may not simply reflect pre-existing biases, but may also provide a distinct source of information for learning about these stereotypes. We refer to this second possibility that language exerts a causal influence on people’s biases as the _language as causal factor_ hypothesis. 

In Study 1, we examine whether language-derived gender biases predict responses on the gender-career IAT. Our analysis focuses on the _distributional_ structure of language rather than the specifics of the communicated content. In Study 2, we examine how the psychological biases measured by the IAT and the linguistic biases we measure relate to more structural aspects of language: sex-based grammatical gender and the prevalence of gender-specific occupation terms (e.g., “waiter”/“waitress” but “teacher”/“teacher”). The results of Study 2 suggest that language not only reflects existing gender biases, but may play a causal role in shaping them. 

```{r child = "sections_pnas/study0_writeup.Rmd"}
```

```{r child = "sections_pnas/study1_writeup.Rmd"}
```

```{r child = "sections_pnas/study2_writeup.Rmd"}
```

# General Discussion
Where do we get our gender stereotypes? Non-linguistic experiences surely play a role, but might we also be learning our biases from the statistics of language to which we are exposed? We used a large-scale dataset of Implicit Association Tests (IATs) measuring the bias to associate men with career and women with family and related people’s measured implicit bias to the statistics of the dominant language spoken in the country of the participants. In Study 1, we found that languages with a greater gender bias in their distributional structure, tend to have speakers that have stronger implicit biases. In Study 2, we found a positive relationship between a structural language feature -- the prevalence of gender-marked occupation terms -- and implicit bias. There is suggestive evidence that this greater implicit bias is mediated by the greater gender bias encoded in the distributional patterns of gender-marked terms.

Our work is the first to characterize the relationship between broad structural patterns in language and cultural stereotypes. The  positive correlation that we find between gender bias in language and gender bias in speakers is consistent both with language playing a causal role in the emergence of cultural stereotypes and the idea that language merely reflects existing stereotypes of its speakers. The positive correlation we find in Study 2 between prevalence of gender-specific terms and implicit bias is most parsimoniously explained by the language-as-causal-factor hypothesis because it is unlikely that language forms change on a timescale that could directly reflect behavior. The two causal forces are not mutually exclusive, and in fact may amplify the effects of each other. Future work could use experimental methods to manipulate language statistics in order to more directly examine these causal influences. 

A central contribution of the current work is that it sheds light onto the potential origins of psychological biases that exist at the level of the individual. Given observed large-scale stuctural patterns of gender inequality, such as differences in STEM participation, researchers from a  range of fields have sought to understand the individual-level causal processes that led to the emergence of structural inequality. But, critically, these previous efforts have taken properties of the indidividual -- such as feelings of self-efficacy in science [@stoet2018gender] and general preferences [@falk2018relationship] -- as largely exogeneous. Here, we  provide a potential explanation of the origins of these psychological biases  by arguing that exposure to biased language statistics could play a causal role in the emergence of these biases at the level of the invididual. Consistent with this account, biases in language statistics are correlated with previous individual-level predictors of STEM inequality, such as self-efficacy in science and general preferences (see SM for details). 

One limitation of our work is the reliance on the IAT, which has been criticized for both its low reliability [@lane2007understanding] and limited external validity [@fazio2003implicit]. Issues of reliability are less relevant here because we use the IAT to measure group-level differences rather than as an individual-difference measure. However, concerns about validity are important particularly because we find that language measures and  explicit psychological measures of gender bias are uncorrelated, though explicit bias was measured in a fairly coarse way. Understanding the full import of linguistic biases on cultural stereotypes would therefore require obtaining measures more closely related to real-world behavior. 

Cultural stereotypes are acquired through experience. Here, we show that group-level differences in implicit bias are strongly correlated with the strength of gender bias encoded in the statistics of different languages. This pattern suggests that the statistics of language use are an important source of cultural experience: The mere process of listening to and producing language exposes one to statistics that may lead to the formation of cultural stereotypes. Many cultural associations present in the statistics of language may be innocuous -- indeed, these statistics may be an important mechanism through which cultural information is transmitted [@lupyan2017wordsascues]. But, in other cases, like the kind of gender stereotypes investigated here, language may play a powerful role in their formation, and ultimately contribute to undesirable real-world consequences like gender inequality in STEM. Understanding  the causal role that language plays in the formation of these stereotypes is therefore an important first step to changing these consequences.


 
\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

---
nocite: | 
  @caliskan2017semantics
  @ciafactbook
  @nosek
  @stoet2018gender
  @miller2015women
  @schoemann2017determining
...


\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
