---
title: GDP control analyses
author: Molly Lewis 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: no
    toc: yes
---
  
******

```{r setup, include = F}

# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(here)
library(lme4)
library(wbstats)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)

```

## Correlations between all measures

gdp_2013 is the measure Bill sent; it is missing three values. gdp_2017 is a measure I obtained.
```{r}
gdp_data <- wb(indicator = "NY.GDP.PCAP.CD", 
               startdate = 2017, 
               enddate = 2017) %>%
  select(iso2c, value) %>%
  rename(gdp_2017 = value)

#wbsearch("gdp per capita")
#NY.GDP.MKTP.CD
#NY.GDP.PCAP.PP.CD

PATH <- "Molly data2.csv"
bill_data <- read_csv(PATH) %>%
  janitor::clean_names()  %>%
  left_join(gdp_data, by = c("country_code" = "iso2c"))  %>%
  mutate_if(is.numeric, scale)
```

```{r corrplot, fig.height = 6}
plot_data <- bill_data %>%
  select_if(is.numeric)

long_corr <- cor(plot_data, 
                use = "pairwise.complete.obs") %>%
  as.data.frame() %>%
  rownames_to_column("v2") %>%
  gather("v1", "estimate", -v2)

long_p <- corrplot::cor.mtest(plot_data, 
                             use = "pairwise.complete.obs")$p %>%
  as.data.frame(row.names = names(plot_data)) %>%
  do(setNames(.,names(plot_data))) %>%
      rownames_to_column("v2") %>%
  gather("v1", "p", -v2)

corr_df <- full_join(long_corr, long_p) %>%
  mutate(estimate_char = case_when(v1 == v2 ~ "", 
                              TRUE ~ as.character(round(estimate,2))),
         estimate = case_when(v1 == v2 ~ as.numeric(NA), 
                              TRUE ~ estimate),
         estimate_color = case_when(p < .05 ~ estimate, TRUE ~ 0 ))

ggplot(corr_df, aes(v1, fct_rev(v2), fill = estimate_color)) + 
  geom_tile() + #rectangles for each correlation
  #add actual correlation value in the rectangle
  geom_text(aes(label = estimate_char), size=3) + 
  scale_fill_gradient2(low ="blue", mid = "white", high = "red", 
                       midpoint = 0, space = "Lab", guide = "colourbar",
                       name = "Pearson's r") +
  ggtitle("Pairwise Correlation Coefficients") +
  theme_classic(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), #, hjust = .95, vjust = .2), 
        axis.title.x=element_blank(), 
        axis.title.y=element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")
```


```{r}
bill_data %>%
  ggplot(aes(x = lang_es_sub, y = women_stem))+
  geom_point(size = 3) +
  geom_smooth(method = "lm", alpha = .2) +
  theme_classic()

bill_data %>%
  ggplot(aes(x = lang_es_sub, y = gdp_2017))+
  geom_point(size = 3) +
  geom_smooth(method = "lm", alpha = .2) +
  theme_classic()


```


## Country-level (subtitles) {.tabset}
### Predicting women in stem with embedding measures, controlling for median age.

Bill's measure:
```{r}
lmer(women_stem ~  lang_es_sub + gdp_2013  + (1|language_name), 
   data = bill_data) %>%
  summary()
```

My measure:
```{r}
lmer(women_stem ~  lang_es_sub + gdp_2017  + (1|language_name), 
   data = bill_data) %>%
  summary()
```

### Predicting women in stem with embedding measure, controlling for median age and gdp.

additive:
```{r}
lmer(women_stem ~  lang_es_sub + median_country_age + gdp_2017 +  (1|language_name), 
   data = bill_data) %>%
  summary()
```

interactive:
```{r}

lmer(women_stem ~  lang_es_sub * median_country_age + gdp_2017 +  (1|language_name), 
   data = bill_data) %>%
  summary()

lmer(women_stem ~  lang_es_wiki * median_country_age + gdp_2017 +  (1|language_name), 
   data = bill_data) %>%
  summary()
```

```{r, include = F}
####  Predicting women in stem with occupation measures, controlling for median age and gdp.

lmer(women_stem ~  subt_occu_semantics_fm*median_country_age + gdp_2017 +
         (1|language_name), 
   data = bill_data) %>%
  summary()

lmer(women_stem ~  mean_prop_distinct_occs * median_country_age  + gdp_2017 +
         (1|language_name), 
   data = bill_data) %>%
  summary()
```

```{r}
mod <- lm(women_stem ~ gdp_2017, 
   data = bill_data) 

bill_data %>%
  modelr::add_residuals(mod, "women_stem_gdp_resid") %>%
  mutate(median_country_age_tile = ntile(median_country_age, 2),
         country_age = ifelse(median_country_age_tile == 1, "younger", "older")) %>%
  ggplot(aes(x = lang_es_sub, y = women_stem_gdp_resid, color = country_age)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", alpha = .2) +
  theme_classic()
```

## Language-level (subtitles) {.tabset}
```{r}
lang_level <- bill_data %>%
  group_by(language_name) %>%
  summarize_if(is.numeric, mean, na.rm = T)

```

### Predicting women in stem with embedding measures, controlling for median age.
```{r}
lm(women_stem ~  lang_es_sub + gdp_2017, 
   data = lang_level) %>%
  summary()

```

### Predicting women in stem with embedding measure, controlling for median age and gdp.

additive:
```{r}
lm(women_stem ~  lang_es_sub + median_country_age + gdp_2017, data = lang_level) %>%
  summary()
```

interactive
```{r}
lm(women_stem ~  lang_es_sub * median_country_age + gdp_2017 , 
   data = lang_level) %>%
  summary()
```

```{r, include = F}
####  Predicting women in stem with occupation measures, controlling for median age and gdp.

lm(women_stem ~  subt_occu_semantics_fm*median_country_age + gdp_2017, 
   data = lang_level) %>%
  summary()

lm(women_stem ~  mean_prop_distinct_occs * median_country_age  + gdp_2017, 
   data = lang_level) %>%
  summary()
```


```{r}

mod <- lm(women_stem ~ gdp_2017, 
   data = lang_level) 

lang_level %>%
  modelr::add_residuals(mod, "women_stem_gdp_resid") %>%
  mutate(median_country_age_tile = ntile(median_country_age, 2),
         country_age = ifelse(median_country_age_tile == 1, "younger", "older")) %>%
  ggplot(aes(x = lang_es_sub, y = women_stem_gdp_resid, color = country_age)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", alpha = .2) +
  theme_classic()

```

```{r}
lang_level %>%
  ggplot(aes(x = lang_es_sub, y = women_stem))+
  geom_point(size = 3) +
  geom_smooth(method = "lm", alpha = .2) +
  theme_classic()

lang_level %>%
  ggplot(aes(x = lang_es_sub, y = gdp_2017))+
  geom_point(size = 3) +
  geom_smooth(method = "lm", alpha = .2) +
  theme_classic()
```

