---
title: IAT demographics by country
subtitle: ED1
date: "`r Sys.Date()`"
output:
  html_document:
  code_folding: hide

---


```{r setup, include = F}
# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(cowplot)
library(here)

opts_chunk$set(echo = F, message = F, warning = F,
               error = F, tidy = F, cache = F)
```

```{r}
full_width = 7.08661
FIG_OUTPATH <- here("writeup/journal/supporting_information/ED/ED1/ED1.pdf")

```

## Write data to supplementary information

```{r}
BY_COUNTRY_DF <- here("data/study0/processed/by_country_df.csv")
SI_OUTPATH <- here("writeup/journal/supporting_information/ED/ED1/ED1_data.csv")
temp <- read_csv(BY_COUNTRY_DF) %>%
  select(country_code, country_name, n_participants, prop_male,median_country_age,age_ci_lower, age_ci_upper, mean_age)
write_csv(temp, SI_OUTPATH)

country_iat <- read_csv(SI_OUTPATH) %>%
  mutate(prop_female = 1 - prop_male)
```


## Plot
```{r}
# participants by country
p1 <- country_iat %>%
  filter(country_name != "United States of America")  %>%
  ggplot(aes(x = fct_reorder(country_name, -n_participants),
             y = n_participants)) +
  geom_bar(stat = "identity")  +
  xlab("") +
  ylab("N participants") +
  scale_y_continuous(breaks = seq(0,30000,5000)) +
  ggtitle("Number of participants by country") +
  #theme_minimal_hgrid(font_size = 7.5) +
  theme_classic(base_size = 7) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,  vjust = .5),
       panel.grid.major.y = element_line(colour = "grey"))

```


```{r}
country_gender <- country_iat %>%
  select(country_name, prop_male) %>%
  rename(Male = prop_male) %>%
  mutate(Female = 1 - Male) %>%
  gather("Gender", "prop", -country_name)

p2 <- country_gender %>%
  ggplot(aes(x = fct_reorder(country_name, -prop), y = prop, fill = Gender, group = Gender)) +
  geom_bar(stat = "identity")  +
  scale_fill_manual(values = c("#D55E00", "#56B4E9")) +
  geom_hline(aes(yintercept = .5), linetype = 2) +
  xlab("") +
  ylab("Proportion participants") +
  ggtitle("Participant gender by country") +
   theme_classic(base_size = 7) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,  vjust = .5),
       panel.grid.major.y = element_line(colour = "grey"),
       legend.key.size = unit(.7,"line"))

```

```{r}
p3 <- country_iat %>%
  ggplot(aes(x = fct_reorder(country_name, -mean_age),
             y = mean_age)) +
  geom_bar(stat = "identity")  +
  geom_point(aes(x = fct_reorder(country_name, -mean_age), y = median_country_age), color = "red", size = .5) +
  geom_linerange(aes(ymin = age_ci_lower, ymax = age_ci_upper)) +
  xlab("") +
  ylab("Mean participant age (years)") +
  ggtitle("Participant age by country") +
  theme_classic(base_size = 7) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,  vjust = .5),
       panel.grid.major.y = element_line(colour = "grey"))


```

```{r}
bottom_row <- plot_grid(p2, p3, labels = c('b', 'c'), label_size = 12, rel_widths = c(1.2, 1))
full_fig <- plot_grid(p1, bottom_row, labels = c('a', ''), label_size = 12, ncol = 1)
full_fig
```

```{r}
pdf(FIG_OUTPATH, width = full_width)
full_fig
dev.off()
```

## Title and Legend 

Sample size and demographic characteristics of Project Implicit data (Study 1b)

a, Number of participants by country after exclusions (note that US participants are excluded from the visualization because of the large sample size; _N_ =  545,673). Our final sample included `r format(sum(country_iat$n_participants), big.mark=",")` participants from `r nrow(country_iat)` countries (see Supplementary Information for exclusion criteria). b, Gender distribution of participants by country after exclusions. Across countries, there tended to be more female participants relative to male participants (_M_ = `r round(mean(country_iat$prop_female),2)` proportion females; _SD_  = `r round(sd(country_iat$prop_female),2)`). c, Age distribution of participants by country after exclusions. Ranges correspond to 95% CIs. Red points show median age by country.


